---
title: "Data Scraping Lab"
author: "Liam Niehus-Staab, Ishaan Tibrewal, Thomas Leistikow"
date: "11/12/2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rvest)
library(stringr)
library(dplyr)
library(mosaic)
library(ggplot2)
```

## Clean Data
```{r} 
baseball = read_html("https://en.wikipedia.org/wiki/Triple_Crown_(baseball)")

#pull out table elements from wiki page
ball = html_nodes(baseball, "table")

#choose pitching table
pitching = html_table(ball, header = TRUE, fill = TRUE)[[3]]

#rename columns
names(pitching)[5:7] = c("EarnedRunAverage", "Wins", "StrikeOuts")

#get rid of asterix in the 3 cols above
pitching$EarnedRunAverage = as.numeric(str_replace_all(pitching$EarnedRunAverage,
                                                       "\\*", ""))
pitching$Wins = as.numeric(str_replace_all(pitching$Wins,
                                                       "\\*", ""))
pitching$StrikeOuts = as.numeric(str_replace_all(pitching$StrikeOuts,
                                                       "\\*", ""))
#remove extra useless refs col
pitching$`Ref(s)` <- NULL

#get batting data
batting = html_table(ball, header = TRUE, fill = TRUE)[[2]]

#delete useless col
batting$`Ref(s)` <- NULL

#rename columns
names(batting)[6:8] = c("HomeRuns", "RunsBattedIn", "BattingAvg")

#get rid of asterix in the 3 cols above
batting$HomeRuns = as.numeric(str_replace_all(batting$HomeRuns,
                                                       "\\*", ""))
batting$RunsBattedIn = as.numeric(str_replace_all(batting$RunsBattedIn,
                                           "\\*", ""))
batting$BattingAvg = as.numeric(str_replace_all(batting$BattingAvg,
                                                 "\\*", ""))
```

## Show Cleaned Data
```{r} 
#show tables
head(batting)
head(pitching)
```

## Plot Data
```{r}
#make plot
ggplot(data = pitching, aes(x = Year, y = Wins)) + 
  geom_point() + stat_smooth(method = lm) + labs(title = "Triple Crown Wins by Year")

```
It can be seen that as time has gon one, winners of the pitching tripple crown
in baseball have had a clearly decreasing ratio of wins. Because having the
triple crown means that each player lead the league in wins, run average, and
strike outs, that means that ever single pitcher in the whole league has been
progressively getting fewer wins to their name as the years go on. This could
say something about teams being more balanced as time goes on, or it may have
something to do with there being more pitchers in each team (causing the wins
to be spread out over a greater number of players) over time. 
