---
title: "Relational Data Lab"
author: "Liam Niehus-Staab"
date: "10/10/2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(dplyr)
library(tidyr)
library(dbplyr)
library(RSQLite)
```
```{r data, include=FALSE, echo=FALSE}
lahman_db <- src_sqlite("~/Shared/F18MAT295/case-studies-master/baseball-salaries/data/lahman.sqlite", create = TRUE)
inflation <- read.csv("~/Shared/F18MAT295/case-studies-master/baseball-salaries/data/inflation.csv")


### Compare payroll across years 

#select salaries table from the db
slaries_tbl = tbl(lahman_db, "salaries")
#get the salaries and payroll
salaries_grouped <- group_by(salaries_tbl, teamID, yearID)
payroll <- summarize(salaries_grouped, payroll = sum(salary, na.rm = TRUE))

#the collect command pulls the db data into R
payroll <- dplyr::collect(payroll)

#join the payroll and inflation data so we can see how payroll is affected by inflation rates
payroll = left_join(payroll, inflation, by = c("yearID" = "year"))
#mutate data to add inflation adjusted payrool col 
payroll <- mutate(payroll, adj.payroll = payroll * multiplier)
teams_tbl <- collect(tbl(lahman_db, "teams"))

#join payroll data with team data
payroll_teams = left_join(payroll, teams_tbl, by = c("teamID", "yearID"))
postseason <- collect(tbl(lahman_db, "seriespost"))
#narrow the data
postseason <- select(postseason, yearID, teamIDwinner, teamIDloser)

#gather the 'teamIDwinner' and 'teamIDloser' columns into 'teamID' and 'win' cols
postseason = gather(postseason, key = win, value = teamID, teamIDwinner, teamIDloser)

#drop the win col since we are not interested in who won 
postseason <- select(postseason,-win)
#now group by team and year so each team only has 1 entry per season
postseason_grouped <- group_by(postseason, yearID, teamID) 

#summarize the data
playoffs <- summarize(postseason_grouped, playoff = "Y")

#merge the payroll data with the playoff data
payroll_postseason = left_join(payroll_teams, postseason, by = c("teamID", "yearID"))
payroll_postseason <- mutate(payroll_postseason, WSWin = ifelse(WSWin == "Y", "Y", "N"))

```

### On Your Own

 The **payroll_postseason** data table created above contains the variables `WCWin` 
and `DivWin` which indicate whether a team won a wild card or divisional playoff
game (`Y`) or not (`N`). 
 a. Use the `mutate` function in the `dplyr` package to fill in `N`s for the blanks
in the `WCWin` and `DivWin` columns. (*Note*: If a team did not reach the playoffs, 
then its entry for `WCWin` and `DivWin` will be blank.)
```{r}
payroll_postseason$DivWin = sub("^$", "N", payroll_postseason$DivWin)
payroll_postseason$WCWin = sub("^$", "N", payroll_postseason$WCWin)
```
 b. Create a dotplot of payroll by year, highlighting teams that won the wild card
round. Briefly summarize your findings.
```{r}
ggplot(data = payroll_postseason, mapping = aes(x = yearID, y = adj.payroll/1000000, color = WCWin)) +
  geom_point(data = filter(payroll_postseason, WCWin == "N")) + 
  geom_point(data = filter(payroll_postseason, WCWin == "Y")) + 
  labs(title = "Wild card round win", x = "Year", y = "Payroll(mils)") + 
  scale_color_manual("Wild Card Round Win", values = c("gray80", "darkorange"))
# Winning the wild card round appears to have no corelation to payroll, the winners come from a 
# wide spread of variously paid teams.
```

 c. Create a dotplot of payroll by year, highlighting teams that won their division. 
Briefly summarize your findings.
```{r}
ggplot(data = payroll_postseason, mapping = aes(x = yearID, y = adj.payroll/1000000, color = DivWin)) +
  geom_point(data = filter(payroll_postseason, DivWin == "N")) + 
  geom_point(data = filter(payroll_postseason, DivWin == "Y")) + 
  labs(title = "Division win", x = "Year", y = "Payroll(mils)") + 
  scale_color_manual("Div Win", values = c("gray80", "darkorange"))
# Having a division win seems to have a mild positive correlation with payroll. However there
# are also many teams that won their division that had low to average payrolls. This
# suggests that the pay of a team may not affect their skill.
```
 d. The distribution of payroll values is quite skewed. One way to adjust for this 
would be to look at a log transformation of team payroll. Remake both graphs from 
the "Payroll and Wins" section (wins v payroll overall, and wins v payroll faceted 
by year) using a natural log transform of the payroll values. Do the general trends
appear the same in the log-transformed data? What features of the data may be easier
to see in the log scale? What features may be obscured by using logs?
```{r}
ggplot(data = payroll_teams, aes(x = log(adj.payroll/1e6), y = W)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  labs(x = "log of payroll (millions)", y = "wins")
# Taking the log of payroll really destroys any shape the data had; the points form a 
# blob. While the tred lines appear similar, the shape of the graph really shows that
# there is no correlation between payroll and wins.

#facet by year to see if trend goes across all seasons
ggplot(data = payroll_teams, aes(x = log(adj.payroll/1e6), y = W)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  facet_wrap(~ yearID, ncol = 5) + 
  labs(x = "log of payroll (millions)", y = "wins")
# Faceting by year shows a similar blobiness. The trend lines appear to all be mostly
# flat.  This provides support for the hypothesis that there is no strong correlation.

# Taking the log of the payroll makes it easier to see that there isn't much of a 
# correlation between wins and pay, as the data appears in blobs. However it makes
# it harder to see that the teams that are paid the most by far do tend to win more
# than average. However, that does not mean that a strong trend exists because teams
# that are paid a lot less also have the same amount or more wins.
```

