---
title: "Terrorism Lab"
author: "Liam Niehus-Staab & Albert"
date: "10/4/2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# load required packages
library(maptools) # creates maps and work with spatial files
library(broom)    # assists with tidy data
library(ggplot2)  # graphics package
library(leaflet)  # interactive graphics (output does not show in RMD files)
library(dplyr)    # joining data frames
library(readr)    # quickly reads files into R

# Reads the csv
TerrorismData <- read_csv("~/Shared/F18MAT295/r-tutorials-master/maps-shapefiles/data/terrorismData.csv")

GTdata <- read.csv("~/Shared/F18MAT295/case-studies-master/Terrorism_CaseStudy/data/GTdata.csv")

Worldshapes <- readShapeSpatial("~/Shared/F18MAT295/r-tutorials-master/maps-shapefiles/data/ne_50m_admin_0_countries")
```
Tidy the Terrorism data
```{r, echo=FALSE}


#show only attacks against government
govAttacks = filter(GTdata, TargetType == "Government")

#Create a new column (with a numeric class) where 1 represents success and 0 represents failure
govAttacks = mutate(govAttacks,
                    SuccessRate=as.numeric(as.logical(GTdata$Success=="successful")))

#Verify that a new column, Success2, is numeric. Numeric columns can be summed.
str(govAttacks)

#Create a table using counts of total incidences, total successful incidences by TargetType, and percent successful
dplyr::summarise(group_by(CBRNIncidents, TargetType), 
                 TotalCount=n(),
                 SuccessfulCount=sum(Success2),
                 PercentSuccessful=SuccessfulCount/TotalCount)



```

Graph 
```{r}
#title
#Success rates of terrism against governments by country
```
Question:
Which governments had the most terrorist attacks against them? How successful were those attacks?

Analysis of Graph:
The graphic shows that Iraq had not only the most total attacks, but also appears to have the greatest number of successful attacks. Nearby Turkmenistan appears to have the lowest attack success rate, although it does not have as many attacks. Some central African countries have relativly high terrorism success rates, but fewer total attacks. Generally, it appears that countries with high terrorism success rates also have a large number of total terrorist attacks. 
