---
title: "CoffeeLab"
author: "Liam Niehus-Staab"
date: "9/18/2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#libs
library(dplyr)
library(ggplot2)
library(mosaic)

# import csv
coffeeSales <- read.csv("~/MAT295/CoffeeLab/coffee.csv")

```


```{r coffeeSales}
#make a graphs
ggplot(data = coffeeSales, aes(x = Location, y = Sales)) + 
  geom_boxplot()  + 
  labs(title = "Box Plot of Sales by Location")

# null Hypothesis: There is no difference in average sales between the 
#                  park and the buisness district
# alternate Hypothesis: There is a non 0 difference between the two locations
business = filter(coffeeSales, Location == "Business")
park = filter(coffeeSales, Location == "Park")
t.test(park$Sales, business$Sales)
```

 Dear Joe,
 My analysis of the sales data you collected is complete. I conducted a simple T test on 
 the data for number of sales by location and found some good evidence that there is a
 difference between the productivity of your shop at the two locations. Now I have some
 good news and some bad news. The good news is that the business district appears to 
 consistently make more sales than the park. The T test showed a very small p value, 
 signifying that the chance that these sales results being pure chance are tiny and gives
 us good reason to reject the hypothesis that there was no difference between the locations. 
 The mean number of sales for the business district is about 14% higher than that of the 
 park. If you wish to continue with your coffee truck, I would recommend selling in the 
 business district. That brings us to the bad news: you didnt make a positive profit
 for any of the days. Really consider if this is what you want to do with your money.


