---
title: "Text Analysis Lab"
author: "Liam Niehus-Staab"
date: "10/29/2018"
output: word_document
---

```{r setup, include=FALSE}
#libs
knitr::opts_chunk$set(echo = TRUE)
library(stringr) #string library
require(stringi, quietly = TRUE,warn.conflicts = FALSE)
# install.packages("knitr")
require(knitr, quietly = TRUE, warn.conflicts = FALSE)
require(ggplot2, quietly = TRUE, warn.conflicts = FALSE) #for qplots

#imports

```
#Intro Strings Questions
```{r}
veggies = c("carrot", "bean", "peas", "cabbage", "scallion", "asparagus")

# + Find those strings that contain the pattern "ea".
str_detect(veggies, "ea") 
```
```{r}
# + Find those strings that end in "s".
str_detect(veggies, "s$")
```
```{r}
# + Find those strings that contain at least two "a"'s.
str_detect(veggies, ".*a.*a")
```
```{r}
# + Find those strings that begin with any letter except "c".
str_detect(veggies, "^[^c]")
```
```{r}
# + Find the starting and ending position of the pattern "ca" in each string.
str_locate_all(veggies, "ca")
```


2. The regular expression `"^[Ss](.*)(t+)(.+)(t+)"` matches "scuttlebutt", "Stetson", and "Scattter", but not "Scatter." Why?
The regex matches either S or s as first char, then any number of any char, then 1 or more t, then 1 or more of any character, then 1 or more t.


It doesn't match "Scatter" because after it matches the first 't' with t+, then the next 't' with .+, and it cant
match the last t+ because there are only 2 't's in Scatter.



##Text Mining
```{r}
#read in blogs, twitter and news
Twitter1 <- readLines("~/Shared/F18MAT295/en_US.twitter.txt",encoding = "UTF-8", skipNul = TRUE, warn = FALSE)

# 1. Are strings involving the words "love" longer or shorter than strings with the word "hate"?
#Strings with love appear to be a little be shorter on average.
```
```{r}
# * How many rows contain the word hate (include upper and lower case letters)?
hate <- sum(grepl("[Hh][Aa][Tt][Ee]", Twitter1))
```
```{r}
# * Give a 5 number summary for the number of characters in texts that include the word hate.
hate2 = grep("[Hh][Aa][Tt][Ee]", Twitter1)
hateTwit = Twitter1[hate2]
summary(nchar(hateTwit))
```
```{r}
# * Give a 5 number summary for the number of characters in texts that include the word love.
love2 = grep("[Ll][Oo][Vv][Ee]", Twitter1)
loveTwit = Twitter1[love2]
summary(nchar(loveTwit))
```
```{r}
# * Create bar charts for both text files.
TwL = stri_count_words(loveTwit) 
qplot(TwL, binwidth = 5, main = "Number of words in each line containing 'love'")
TwH = stri_count_words(hateTwit) 
qplot(TwH, binwidth = 5, main = "Number of words in each line containing 'hate'")

```